\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-09-03 smurf_paper.tex
Update common mode section

The \model{COM} model is the common-mode signal itself. It is a single
time series that is estimated by finding the mean of all bolometers at
each time step. Bolometer values that have been flagged as unusable
for any reason are excluded from the mean. {\bf Per: has a comment in
here about stability of the gain pattern... it should be constant
right?}.

Since bolometers have varying sensitivities, the amplitude of the
common-mode variations will also vary from bolometer to bolometer. So
comparing each bolometer time stream with the common-mode signal
allows an estimate of the bolometer sensitivity to be obtained. In
practice, a least squares linear fit is performed between the
bolometer time series and the common-mode to determine a gain and
offset for each bolometer.  The gain and offset for each bolometer is
known as the \model{GAI} model. Each bolometer value can then
optionally be scaled and shifted using these values so that all
bolometers share a common (but as yet unknown) calibration. This
provides an alternative (or additional) flat-fielding strategy to that
described in section~\ref{sec:bolos}.

@@

It can be seen from Fig.~\ref{fig:com} that some bolometers depart
radically from the common-mode, indicating some problem within the
bolometer. Such bolometers are identified by calculating the Pearson
coefficient of correlation between each bolometer time-stream and the
common mode. Bolometers for which the correlation coefficient is below
a specified limit, or which have a negative gain, are flagged as bad
in order to omit them from the final map. If the option described
above for handling time-varying sensitivities is used, then a
correlation coefficient can be determined for each individual block of
time slices.  This allows individual bad blocks to be rejected from a
bolometer time stream, rather than rejecting the whole bolometer. This
is beneficial for bolometers that follow the common-mode for most of
the time but suffer from transient problems at specific times. {\bf
Per: If magnetic pickup dominates the common mode a negative gain is
normal and should not be removed. This might be a risk at high
elevation in very good stable weather!}

The common-mode is calculated from all bolometers that have not
previously been flagged as bad. Thus, aberrant bolometers with low
correlation coefficients will be included, polluting the common-mode. For
this reason, an iterative approach is taken to estimating the common-mode.
The initial common-mode estimate is formed from all remaining bolometers.
Aberrant bolometers with low correlation coefficients are then identified
by comparing each bolometer time stream with the common-mode. Such
aberrant bolometers are flagged as bad, and a new estimate of the
common-mode is produced omitting the newly flagged bolometers. This
process is repeated until no further bolometers are rejected.

>>>>

The \model{COM} model is the common-mode signal itself. It is a single
time series that is estimated by finding the mean of all bolometers at
each time step. Bolometer values that have been flagged as unusable
are excluded from the mean. {\bf Per: has a comment in
here about stability of the gain pattern... it should be constant
right?}.

Even after flat-fielding, bolometers may have slightly varying
sensitivities and so the amplitude of the common-mode variations will
also vary from bolometer to bolometer. Comparing each bolometer time
stream with the common-mode signal allows an estimate of the relative
bolometer sensitivity to be obtained. In practice, a least squares linear
fit is performed between the bolometer time series and the common-mode to
determine a gain and offset for each bolometer.  The gain and offset for
each bolometer is known as the \model{GAI} model. Each bolometer value
can then optionally be scaled and shifted using these values so that all
bolometers share a common (but as yet unknown) calibration. This provides
an alternative (or additional) flat-fielding strategy to that described
in section~\ref{sec:bolos}.

@@

It can be seen from Fig.~\ref{fig:com} that some bolometers depart
radically from the common-mode, indicating some problem within the
bolometer. Such bolometers are identified by calculating the Pearson
coefficient of correlation between each bolometer time-stream and the
common mode. Bolometers for which the correlation coefficient is below
a specified limit, or which have unusually high or low gain compared to
the other bolometers, are flagged as bad in order to omit them from the
final map. If the option described
above for handling time-varying sensitivities is used, then a
correlation coefficient can be determined for each individual block of
time slices.  This allows individual bad blocks to be rejected from a
bolometer time stream, rather than rejecting the whole bolometer. This
is beneficial for bolometers that follow the common-mode for most of
the time but suffer from transient problems at specific times. {\bf
Per: If magnetic pickup dominates the common mode a negative gain is
normal and should not be removed. This might be a risk at high
elevation in very good stable weather!}

The common-mode value at each time slice is calculated as the unweighted
mean of the values of all bolometers that have not previously been
flagged as bad. Thus, aberrant bolometers with low correlation
coefficients will be included, polluting the common-mode. By default, no
attempt is made to exclude these aberrant bolometers\footnote{An option
exists to reject unusually high or low bolometer values when calculating
the common mode value at each time slice, but this seems to introduce
instability into the convergence process and makes little difference to
the final maps.}.

Bolometer values that were flagged as poorly correlated with the common
mode on the previous iteration are nevertheless included in the
calculation of the common mode on the next iteration. If sensitivities
are allowed to vary with time as described above, then poorly correlated
values will be flagged in blocks of - typically of 30 seconds duration.
Excluding such flagged blocks from subsequent estimates of the common
mode can result in discontinuities in the common mode signal at the
edges of these blocks. This in turn can lead to ringing in the low pass
filter, and instabilities in the convergence process.
